\textbf{SPECFEM3D Cartesian}\\ \textbf{User Manual}

### 2014-10-03 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
replaced latitude with latorUTM and longitude with longorUTM in CMTSOLUTION and FORCESOLUTION (as mentioned by Alexis Bottero the original names did not make sense in the cause of SUPPRESS_UTM_PROJECTION = .true. because the values are then given in UTM rather than longitude and latitude). The code still accepts the old format for back-compatibility reasons. That is also why I kept the same length as before for the new names.
Adds note on MPI-IO

Then, to configure the software for your system, run the \texttt{configure}
shell script. This script will attempt to guess the appropriate configuration
values for your system. However, at a minimum, it is recommended that
you explicitly specify the appropriate command names for your Fortran
compiler and MPI package (another option is to define FC, CC and MPIF90
in your .bash\_profile or your .cshrc file):
\begin{lyxcode}
./configure~FC=ifort~MPIFC=mpif90
\end{lyxcode}
Before running the \texttt{configure} script, you should probably
edit file \texttt{flags.guess} to make sure that it contains the best
compiler options for your system. Known issues or things to check
are:
\begin{description}
\item [{\texttt{Intel ifort compiler}}] See if you need to add \texttt{-assume byterecl} for your machine. \textbf{In the case of that compiler, we have noticed that versions dot zero sometimes have bugs or issues that can lead to wrong results when running the code, thus we \emph{strongly} recommend using versions dot one or above (for instance version 13.1 instead of 13.0, version 14.1 instead of 14.0 and so on)}.
\item [{\texttt{IBM compiler}}] See if you need to add \texttt{-qsave}
or \texttt{-qnosave} for your machine.
\item [{\texttt{Mac OS}}] You will probably need to install \texttt{XCODE}.
In addition, the \texttt{clock\_gettime} routine, which is used by
the \texttt{SCOTCH} library that we use, does not exist in Mac OS.
You may need to replace it with \texttt{clock\_get\_time} if you
want to use \texttt{SCOTCH}.
\end{description}
When compiling on an IBM machine with the \texttt{xlf} and \texttt{xlc}
compilers, we suggest running the \texttt{configure} script with the
following options:\\

@@

\end{lyxcode}
The \texttt{CMTSOLUTION} file should be edited in the following way:
\begin{itemize}
\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
will not run otherwise.) The time shift parameter would simply apply
an overall time shift to the synthetics, something that can be done
in the post-processing (see Section \ref{sec:Process-data-and-syn}).
\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
we recommend setting the source half-duration parameter \texttt{half
duration} equal to zero, which corresponds to simulating a step source-time
function, i.e., a moment-rate function that is a delta function. If
\texttt{half duration} is not set to zero, the code will use a Gaussian
(i.e., a signal with a shape similar to a `smoothed triangle', as
explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
source-time function with half-width \texttt{half duration}. We prefer
to run the solver with \texttt{half duration} set to zero and convolve
the resulting synthetic seismograms in post-processing after the run,
because this way it is easy to use a variety of source-time functions
(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
that the noise generated in the simulation by using a step source
time function may be safely filtered out afterward based upon a convolution
with the desired source time function and/or low-pass filtering. Use
the serial code \texttt{convolve\_source\_timefunction.f90} and the
script \texttt{convolve\_source\_timefunction.csh} for this purpose,
or alternatively use signal-processing software packages such as SAC
\urlwithparentheses{www.llnl.gov/sac}. Type

@@

\noindent The \texttt{FORCESOLUTION} file should be edited in the
following way:
\begin{itemize}
\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
will not run otherwise.) The time shift parameter would simply apply
an overall time shift to the synthetics, something that can be done
in the post-processing (see Section \ref{sec:Process-data-and-syn}).
\item Set the \texttt{half duration} parameter of the Ricker source time
function when {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned on
in the main parameter file {\texttt{Par\_file}}. In case that the
solver uses a (pseudo) Dirac delta source time function to represent
a force point source, a very short \texttt{half duration} parameter
(hdur = 5 {*} DT) is automatically set by default.
\item Set the latitude, longitude, depth of the source in geographical coordinates.
\item Set the magnitude of the force source.
\item Set the components of a (non-unitary) direction vector for the force
source in the East/North/Vertical basis (see Appendix A for the orientation
of the reference frame).
\end{itemize}
\noindent Where necessary, set a \texttt{FORCESOLUTION} file in the
same way you configure a \texttt{CMTSOLUTION} file with $N_{\mathrm{sources}}$
entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
\texttt{FORCESOLUTION} files to a single \texttt{FORCESOLUTION} file).
At least one entry (not necessarily the first) must have a zero \texttt{time
shift}, and all the other entries must have non-negative \texttt{time
shift}. Each subevent can have its own half latitude, longitude, depth,
\texttt{half duration} and force parameters.

>>>>

Then, to configure the software for your system, run the \texttt{configure}
shell script. This script will attempt to guess the appropriate configuration
values for your system. However, at a minimum, it is recommended that
you explicitly specify the appropriate command names for your Fortran
compiler and MPI package (another option is to define FC, CC and MPIF90
in your .bash\_profile or your .cshrc file):
\begin{lyxcode}
./configure~FC=ifort~MPIFC=mpif90
\end{lyxcode}
Note that MPI should have been installed with MPI-IO enabled.

Before running the \texttt{configure} script, you should probably
edit file \texttt{flags.guess} to make sure that it contains the best
compiler options for your system. Known issues or things to check
are:
\begin{description}
\item [{\texttt{Intel ifort compiler}}] See if you need to add \texttt{-assume byterecl} for your machine. \textbf{In the case of that compiler, we have noticed that versions dot zero sometimes have bugs or issues that can lead to wrong results when running the code, thus we \emph{strongly} recommend using versions dot one or above (for instance version 13.1 instead of 13.0, version 14.1 instead of 14.0 and so on)}.
\item [{\texttt{IBM compiler}}] See if you need to add \texttt{-qsave}
or \texttt{-qnosave} for your machine.
\item [{\texttt{Mac OS}}] You will probably need to install \texttt{XCODE}.
In addition, the \texttt{clock\_gettime} routine, which is used by
the \texttt{SCOTCH} library that we use, does not exist in Mac OS.
You may need to replace it with \texttt{clock\_get\_time} if you
want to use \texttt{SCOTCH}.
\end{description}
When compiling on an IBM machine with the \texttt{xlf} and \texttt{xlc}
compilers, we suggest running the \texttt{configure} script with the
following options:\\

@@

\end{lyxcode}
The \texttt{CMTSOLUTION} file should be edited in the following way:
\begin{itemize}
\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
will not run otherwise.) The time shift parameter would simply apply
an overall time shift to the synthetics, something that can be done
in the post-processing (see Section \ref{sec:Process-data-and-syn}).
\item For point-source simulations (see finite sources, page \pageref{To-simulate-a})
we recommend setting the source half-duration parameter \texttt{half
duration} equal to zero, which corresponds to simulating a step source-time
function, i.e., a moment-rate function that is a delta function. If
\texttt{half duration} is not set to zero, the code will use a Gaussian
(i.e., a signal with a shape similar to a `smoothed triangle', as
explained in \citet{KoTr02a} and shown in Fig~\ref{fig:gauss.vs.triangle})
source-time function with half-width \texttt{half duration}. We prefer
to run the solver with \texttt{half duration} set to zero and convolve
the resulting synthetic seismograms in post-processing after the run,
because this way it is easy to use a variety of source-time functions
(see Section \ref{sec:Process-data-and-syn}). \citet{KoTr02a} determined
that the noise generated in the simulation by using a step source
time function may be safely filtered out afterward based upon a convolution
with the desired source time function and/or low-pass filtering. Use
the serial code \texttt{convolve\_source\_timefunction.f90} and the
script \texttt{convolve\_source\_timefunction.csh} for this purpose,
or alternatively use signal-processing software packages such as SAC
\urlwithparentheses{www.llnl.gov/sac}. Type

@@

\noindent The \texttt{FORCESOLUTION} file should be edited in the
following way:
\begin{itemize}
\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
will not run otherwise.) The time shift parameter would simply apply
an overall time shift to the synthetics, something that can be done
in the post-processing (see Section \ref{sec:Process-data-and-syn}).
\item Set the \texttt{half duration} parameter of the Ricker source time
function when {\texttt{USE\_RICKER\_TIME\_FUNCTION}} is turned on
in the main parameter file {\texttt{Par\_file}}. In case that the
solver uses a (pseudo) Dirac delta source time function to represent
a force point source, a very short \texttt{half duration} parameter
(hdur = 5 {*} DT) is automatically set by default.
\item Set the latitude or UTM $x$ coordinate, longitude or UTM $y$ coordinate, depth of the source (in km).
\item Set the magnitude of the force source.
\item Set the components of a (non-unitary) direction vector for the force
source in the East/North/Vertical basis (see Appendix A for the orientation
of the reference frame).
\end{itemize}
\noindent Where necessary, set a \texttt{FORCESOLUTION} file in the
same way you configure a \texttt{CMTSOLUTION} file with $N_{\mathrm{sources}}$
entries, one for each subevent (i.e., concatenate $N_{\mathrm{sources}}$
\texttt{FORCESOLUTION} files to a single \texttt{FORCESOLUTION} file).
At least one entry (not necessarily the first) must have a zero \texttt{time
shift}, and all the other entries must have non-negative \texttt{time
shift}. Each subevent can have its own half latitude, longitude, depth,
\texttt{half duration} and force parameters.
